A Method for the Spectroscopic Inference of Stellar Parameters

### 2014-07-06 tex/paper.tex
mor eburn in

\subsubsection{Global Covariance Structure}

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{correlation_panel.pdf}
  \caption{\textbf{Top}: the big spectrum. \textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.   
  \textbf{Right} The autocorrelation (Equation~\ref{eqn:autocorrelation}) of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
\label{fig:class0}
\end{center}
\end{figure*}

@@

For a well-designed spectrograph and sufficiently accurate model spectrum, this
{\it global} (i.e., present throughout the spectrum) covariant structure should
have a relatively low amplitude and small correlation length.  To describe that
structure, we assume a stationary covariance kernel (also called a radial basis
function) with an amplitude that depends only on the velocity distance between
two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
between pixel residuals, such that
\begin{equation}
  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian
processes to parameterize such a mildly covariant structure
\citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
$\nu = 3/2$,
\begin{equation}
  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
\end{equation}
which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel (so the effective kernel is $w \, \mathcal{K}_g$).  The truncation distance 
$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
structure in the residual spectrum that is similar to those seen in a typical data--model 
comparison.

@@

\subsubsection{Local Covariance Structure}

@@

These local kernels can be further modified to account for more complex residual structures.  For 
example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
spectral scale.  This phenomenologically different local covariance behavior can still be treated 
in this framework, if we permit the kernel in Eq.~\ref{eqn:klocal} to be modified by another 
(stationary) function.  Here we assume that function is a squared exponential,
\begin{equation}
  \mathcal{S}_k(\lambda_i, \lambda_j |\, h_k) = \exp \left ( - \, \frac{r_{ij}^2}{2\,h_k^2}\right ),
\end{equation}
where $h_k$ is a bandwidth parameter; small $h_k$ generates high-frequency structure, and vice 
versa.  The functionality of this kernel modification (where the appropriate kernel is now $w_k \, 
\mathcal{S}_k \, \mathcal{K}_k$) is demonstrated for the worked example presented in Section 
\ref{subsec:mdwarf}.  
\comm{I have to admit that I do not really understand this very well.  I think some appropriate 
figures in Sect 3.2 will really help, though, so I'll wait for that before decided whether or not 
we need to change this part of the text.}

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:line_total}). The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals. These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix. The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix. The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
\label{fig:region}
\end{center}
\end{figure*}

@@

When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals. Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered. In a single order of an echelle spectrum, there may be $N$ regions of high covariance that are parameterized by several line kernels (Equation~\ref{eqn:line}), which we group into an aggregate parameter $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$. Along with the global covariance parameters and Chebyshev parameters for this order, we call the collection of nuisance hyperparameters for a specific order  (e.g., order $1$) $\vtorder{1} = \{\vtcheb, \vtglobal, \vtlines\}$. Taken together, the aggregated nuisance parameters for all of the $N$ orders are stored in $\vtorders = \{\vtorder{1}, \vtorder{2}, \ldots, \vtorder{N} \}$.

@@

The traditional practice method of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.  Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight. A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit. The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters. In fact, it may be the case that the lines that are off by a small amount (class 0 and III lines in Figure~\ref{fig:badlines}) actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.

Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch. When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length. This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result. However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit.

The benefits of using covariance hyperparameters extend beyond the use case of fitting a single stellar spectrum. If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).

We use a Markov chain Monte Carlo (MCMC) algorithm coupled to a Gibbs sampler to explore the posterior distribution (Equation~\ref{eqn:lnprob}). The MCMC routine efficiently explores a high dimensional probability space using a stochastic, iterative approach. The Gibbs sampler provides a way to easily sample a large number of parameters in a simple, organized manner by sampling only a subset of parameters at any given time, but then rotating among all of the subsets of parameters. For more on MCMC and Gibbs samplers see \citet[ch 11]{gelman13} and the references therein.

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler. These stellar parameters are marginalized over the Chebyshev and noise parameters. Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
}
\label{fig:stellar_posterior}
\end{center}
\end{figure*}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=5in]{figs/stellar_triangle_WASP14_Kurucz.png}
  \caption{The posterior probability function of the stellar parameters for WASP-14, logg fixed to match \citep{torres12}. Using three orders, Kurucz grid. Same assumptions.}
\label{fig:stellar_posterior_logg}
\end{center}
\end{figure*}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters. \textbf{Right}: The posterior probability function of the line parameters, marginalized over the stellar and global noise parameters.}
\label{fig:noise_posterior}
\end{center}
\end{figure*}

@@

\begin{enumerate}
  \item Sample in the stellar parameters. For each proposal of the Metropolis-Hastings algorithm, generate a model spectrum following the steps in \S\ref{subsec:postprocess}. When deciding whether or not to accept the parameters, the Gibbs sampler evaluates $p(\vtstar^i | \vtorders^{i-1})$.
  \item For each order of the echelle spectrum
    \begin{enumerate}
      \item Sample in the Chebyshev polynomial parameters. Adjust the spectrum as detailed in \S\ref{subsec:postprocess}.Gibbs sampler evaluates $p(\vtcheb^{i} | \vtstar^{i}, \vtglobal^{i -1}, \vtlines^{i-1})$.
      \item Sample in the global covariance parameters. Adjust the covariance matrix $C$ as described in \S\ref{subsec:covariance}. Gibbs sampler evaluates $p(\vtglobal^{i}| \vtstar^{i}, \vtcheb^{i}, \vtlines^{i -1})$.

      \item Check to see whether the algorithm should instantiate/delete new/old line kernels
      \item For each line kernel $k$
	\begin{enumerate}
	  \item Sample in the parameters for each line, conditional on the parameters for all the other lines. Adjust the covariance matrix $C$ as described in \S\ref{subsec:covariance}. Gibbs sampler evaluates $p(\vtline{k}^i| \vt_{\textrm{lines}\ne k}^{i - 1})$.
	\end{enumerate}
    \end{enumerate}
\end{enumerate}

At each stage of the Gibbs sampler, the likelihood function (Equation~\ref{eqn:lnprob}) is evaluated for a proposal of a particular subset of parameters, conditional on the current values of all the other parameters. This hierarchical parameter structure and the kernel parameterization of the covariance matrix influences how the algorithm may be efficiently implemented in code. For a typical optical spectrum with $\gtrsim$1,000 pixels, the most computationally intensive step of the likelihood evaluation is usually the matrix product $\fR^T C^{-1} \fR$. Since we have designed the covariance matrix to be sparse, we can use optimized sparse matrix algorithms which are much faster and memory efficient than dense matrix operations. Because we are not interested in the matrix inverse $C^{-1}$ by itself, but rather the product with the residual vectors, we can use efficient routines for solving linear systems to bypass the computationally difficult step of matrix inversion. Additionally, because the covariance matrix is positive semi-definite, we can use the Cholesky factorization of the matrix to optimize the evaluation of the matrix product. Once the covariance matrix is factorized, any subsequent evaluation of the matrix product for different residual vectors $\fR$ is extremely rapid. This makes the $\vtstar$ and $\vtcheb$ steps of the Gibbs sampler extremely fast.
When we sample in the nuisance parameters which affect the covariance matrix, we must redo the Cholesky factorization of $C$ for each update. However, because we designed the kernels to deliver a sparse matrix these operations are efficient enough to be used with an MCMC algorithm.  We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{chen08, davis09} and the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{foreman-mackey12}.

@@

The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters. If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and then the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed. This leads to estimates of stellar parameters with more uncertainty than necessary. 
\todo{Since this point is better carried with an example or test using the global covariance structure and not using it (see Test 1), we might want to defer this discussion to the testing section.}

@@

Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct defects in the synthetic models. \todo{more delicate phrasing?} Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution. This fundamental application of machine learning would enable us to create our own library of semi-empirical stellar models. Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.

\section{Worked Examples}
\label{sec:examples}
Here are some ideas of which tests we might want to show.

@@

[This may need to be eventually moved to \S2, but here it is for now] Cite papers that acquired the data. Describe the instruments. Describe when it was taken. Describe the noise properties, etc. Maybe a table might be worthwhile to describe the spectra. Here also describe the WASP-14 fit using the Kurucz spectra, using a fixed log g, and state that they are the same as Willie's paper. Talk about putting things in a Bayesian framework, using a sensitivity to a prior, to a $\delta$-function prior. Point out that this Bayesian approach allows you to do this.

@@

Talk about giving more reasonable errors using regions, in the biased case.

>>>>

\subsubsection{Global Covariance Structure}
\label{subsec:global_covariance}

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics{correlation_panel.pdf}
  \caption{\textbf{Top}: the big spectrum. \textbf{Left} The same low-amplitude, mildly covariant residuals from Figure~\ref{fig:badlines}, panel 0, but enlarged to show the mildly covariant structure produced by slight mismatch between the data and model spectra.   
  \textbf{Right} The autocorrelation of the residual sequence shown at left. Notice that there is significant correlation for offsets of $\lesssim 4$ pixels.}
\label{fig:class0}
\end{center}
\end{figure*}

@@

For a well-designed spectrograph and sufficiently accurate model spectrum, this
{\it global} (i.e., present throughout the spectrum) covariant structure should
have a relatively low amplitude and small correlation length.  To describe that
structure, we assume a stationary covariance kernel (also called a radial basis
function) with an amplitude that depends only on the velocity distance between
two pixels, 
\begin{equation}
  r_{ij} = r(\lambda_i, \lambda_j) = \Delta v = \frac{c}{2} \left | \frac{\lambda_i 
   - \lambda_j}{ \lambda_i + \lambda_j} \right |,
\end{equation}
where $c$ is the speed of light.  The kernel is chosen to parametrically describe the covariance 
between pixel residuals, such that
\begin{equation}
  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j) =  \langle \vR_i \; \vR_j \rangle.
  \label{eqn:expectation}
\end{equation}
A variety of kernels have been used in the field of Gaussian
processes to parameterize such a mildly covariant structure
\citep[e.g.,][]{rasmussen05}.  Here we adopt the \matern\ kernel with 
$\nu = 3/2$,
\begin{equation}
  \mathcal{K}_{\rm g}(\lambda_i, \lambda_j \, |\, a_{\rm g},\, \ell) = a_{\rm g} \left(1 + \frac{\sqrt{3}\, r_{ij}}{\ell} \right ) \exp 
   \left (- \frac{\sqrt{3}\, r_{ij}}{\ell} \right ),
   \label{eqn:global}
\end{equation}
which is parameterized by an amplitude $a_g$ and scale $\ell$ and makes for a smooth transition to 
negligible covariance at large $r$.  To ensure that $\vC$ remains a relatively sparse matrix that 
enables computational expendiency, we employ a Hann window function
\begin{equation}
  w\,(\lambda_i, \lambda_j |\, r_0) = \left \{ 
    \begin{array}{cc}
    \frac{1}{2} + \frac{1}{2} \cos \left(\frac{\pi r_{ij}}{r_0} \right) & r_{ij} \le r_0 \\
    0 & r_{ij} > r_0 \\
  \end{array}
  \right .
  \label{eqn:Hann}
\end{equation}
to taper the kernel (so the effective kernel is $w \, \mathcal{K}_g$).  The truncation distance 
$r_0$ can be fixed to some reasonable multiple of the scale parameter (here we set $r_0 = 4\ell$).  
The examples in Figure \ref{fig:matern} demonstrate that such a kernel readily produces correlated 
structure in the residual spectrum that is similar to those seen in a typical data--model 
comparison.

@@

\subsubsection{Local Covariance Structure}
\label{subsec:local_covariance}

@@

These local kernels can be further modified to account for more complex residual structures.  For 
example, late-type stars with imperfectly modeled molecular bandheads may produce a complicated 
pattern of positive and negative residuals or a pronounced mismatch over a relatively large 
spectral scale.  This phenomenologically different local covariance behavior can still be treated 
in this framework, if we design an appropriate covariance kernel. Given the behavior of molecular bandheads, a \matern\ kernel tapered by the Gaussian kernel in Eq.~\ref{eqn:klocal} works well. 
\todo{We'll need to round out the Gl 51 example in Sect 3.2 before this text is finished.}

@@

\begin{figure*}[!htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{figs/gauss_matrix.pdf}
\includegraphics[width=0.4\textwidth]{figs/gauss_draw.pdf}
\caption{\textbf{Left} A typical covariance matrix including the Gaussian line kernel (Equation~\ref{eqn:klocal}). The same global covariance shown in Figure~\ref{fig:matern} is still present with the same hyperparameters, however now there is an additional patch of high covariance corresponding to the large, Gaussian-shaped residuals. These larger elements in the covariance matrix effectively down-weight the contribution of a poorly modeled synthetic spectral line.
\textbf{Right} The same spectroscopic residuals shown in Figure~\ref{fig:badlines}, class I, shown with three random draws from the covariance matrix. The random draws can take on a range of amplitudes--positive, negative, or even flat--because they are simply random draws that are described by the covariance matrix. The wide range of possible residual amplitudes match the structure and amplitude of the pixel residuals.}
\label{fig:region}
\end{center}
\end{figure*}

@@

When iteratively fitting a spectrum (see \S\ref{subsec:MCMC}), we continue to add line covariance kernels until we have covered all of the high amplitude residuals. Typically, we will add line kernels until all residuals greater than three times the amplitude of the global covariance kernel are covered. In a single order of an echelle spectrum, there may be $N$ regions of high covariance that are parameterized by several line kernels (Equation~\ref{eqn:klocal}), which we group into an aggregate parameter $\vtlines = \{\vtline{1}, \vtline{2}, \ldots, \vtline{N}\}$. Along with the global covariance parameters and Chebyshev parameters for this order, we call the collection of nuisance hyperparameters for a specific order  (e.g., order $1$) $\vtorder{1} = \{\vtcheb, \vtglobal, \vtlines\}$. Taken together, the aggregated nuisance parameters for all of the $N$ orders are stored in $\vtorders = \{\vtorder{1}, \vtorder{2}, \ldots, \vtorder{N} \}$.

@@

The traditional manner of dealing with spectral mismatch is to simply mask out the regions of the spectrum which do not agree to within a certain tolerance.  Rather than arbitrarily excluding regions of the spectrum from the fit, these regions should instead be incorporated into in the fit with the appropriate weight. A model that includes covariance is far more likely than forcing the synthetic spectrum to fit perfectly, and far more flexible than arbitrarily masking regions which do not fit. The fitting procedure allows these weights to be determined self-consistently, such that lines which are slightly wrong can still bring information to bear on the stellar parameters. In fact, for some types of stars with difficult-to-model spectra, it may be possible that the lines that are off by a small amount (class 0 and III lines in Figure~\ref{fig:badlines}) actually provide the \emph{most information} about the stellar parameters, precisely because these lines are the most sensitive to stellar structure and consequentially are the most difficult to model correctly.

Another powerful benefit of including the covariance hyperparameters is the result of their ability to quantify and account for model-data mismatch. When a data spectrum is fit with a high-quality spectral library, the pixel residuals are less likely to be correlated, and the covariance structure will have a smaller amplitude and correlation length. This means that the inference on the stellar parameters will be more precise, in the same way that high quality data allows a more precise result. However, if the systematic mismatch between the data and model is large, then the hyperparameters will be larger and the precision of the stellar parameters will naturally inflate to respect the quality of model-data fit. The benefits of using covariance hyperparameters also extend beyond the use case of fitting a single stellar spectrum. If we fit many stars with the same set of synthetic models, we can use the structure of the covariance matrix to improve the models themselves (see \S\ref{subsec:learning}).

We use a Markov chain Monte Carlo (MCMC) algorithm coupled to a Gibbs sampler to explore the posterior distribution (Equation~\ref{eqn:lnprob}). The MCMC routine efficiently explores a high dimensional probability space using a stochastic, iterative approach. The Gibbs sampler provides a way to easily sample a large number of parameters in a simple, organized manner by sampling only a subset of parameters at any given time, but then rotating among all of the subsets of parameters. For more on MCMC and Gibbs samplers see \citet[Ch. 11]{gelman13} and the references therein.

@@

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[width=5in]{figs/stellar_triangle_Kurucz.png}
  \caption{The posterior probability function of the stellar parameters for WASP-14, an F star, as explored by the MCMC Gibbs sampler. These stellar parameters are marginalized over the Chebyshev ($\vT_{\rm cheb}$) and noise parameters ($\vT_{\rm cov}$). Figure generated with \texttt{triangle.py} \citep{foreman-mackey14}.
}
\label{fig:stellar_posterior}
\end{center}
\end{figure*}

\begin{figure*}[!htb]
\begin{center}
  \includegraphics[draft, width=3in, height=3in]{figs/global_posteriors.pdf}
  \includegraphics[draft, width=3in, height=3in]{figs/line_posteriors.pdf}
  \caption{\textbf{Left}: The posterior probability function of the global covariance parameters, as explored by the MCMC Gibbs sampler, marginalized over the stellar and line noise parameters. \textbf{Right}: The posterior probability function of the parameters for a single line kernel, marginalized over the stellar and global noise parameters.}
\label{fig:noise_posterior}
\end{center}
\end{figure*}

@@

\begin{enumerate}
  \item Sample in the stellar parameters. For each proposal of the Metropolis-Hastings algorithm, generate a model spectrum following the steps in \S\ref{subsec:postprocess}. When deciding whether or not to accept the parameters, the Gibbs sampler evaluates $p(\vtstar^i | \vtorders^{i-1})$.
  \item For each order of the echelle spectrum
    \begin{enumerate}
      \item Sample in the Chebyshev polynomial parameters. Adjust the spectrum as detailed in \S\ref{subsec:postprocess}.Gibbs sampler evaluates $p(\vtcheb^{i} | \vtstar^{i}, \vtglobal^{i -1}, \vtlines^{i-1})$.
      \item Sample in the global covariance parameters. Adjust the covariance matrix $C$ as described in \S\ref{subsec:global_covariance}. Gibbs sampler evaluates $p(\vtglobal^{i}| \vtstar^{i}, \vtcheb^{i}, \vtlines^{i -1})$.

      \item Check to see whether the algorithm should instantiate/delete new/old line kernels
      \item For each line kernel $k$
	\begin{enumerate}
	  \item Sample in the parameters for each line, conditional on the parameters for all the other lines. Adjust the covariance matrix $C$ as described in \S\ref{subsec:local_covariance}. Gibbs sampler evaluates $p(\vtline{k}^i| \vt_{\textrm{lines}\ne k}^{i - 1})$.
	\end{enumerate}
    \end{enumerate}
\end{enumerate}

At each stage of the Gibbs sampler, the likelihood function (Equation~\ref{eqn:lnprob}) is evaluated for a proposal of a particular subset of parameters, conditional on the current values of all the other parameters. The hierarchical structure of the parameters influences how the algorithm may be efficiently implemented in code. For a typical optical spectrum with $\gtrsim$1,000 pixels, the most computationally intensive step of the likelihood evaluation is usually the matrix product $\fR^T C^{-1} \fR$. Since we have designed the covariance matrix to be sparse, we can use optimized sparse matrix algorithms which are much faster and memory efficient than dense matrix operations. Because we are not interested in the matrix inverse $C^{-1}$ by itself, but rather the product with the residual vectors, we can use efficient routines for solving linear systems to bypass the computationally difficult step of matrix inversion. Additionally, because the covariance matrix is positive semi-definite, we can use the Cholesky factorization of the matrix to optimize the evaluation of the matrix product. Once the covariance matrix is factorized, any subsequent evaluation of the matrix product for different residual vectors $\fR$ is extremely rapid. This makes the $\vtstar$ and $\vtcheb$ steps of the Gibbs sampler extremely fast.
When we sample in the nuisance parameters which affect the covariance matrix, we must redo the Cholesky factorization of $C$ for each update. However, because we designed the kernels to deliver a sparse matrix these operations are still moderately efficient.  We use the high-performance \texttt{SuiteSparse/CHOLMOD}\footnote{\url{http://www.cise.ufl.edu/research/sparse/cholmod/}} library to implement the sparse matrix and Cholesky factorization operations \citep{chen08, davis09} and utilize the Metropolis-Hastings sampler included in the \texttt{Python} MCMC package \texttt{emcee} \citep{foreman-mackey12}.

@@

The benefit of including nuisance parameters and then marginalizing over them is that we can self-consistently model the uncertainty inherent to any spectrum while naturally capturing any degeneracy between the model parameters. If stellar parameters are estimated using a method which ignores these nuisance parameters (e.g., by-eye fitting) and the astronomer arbitrarily inflates the parameter uncertainties to reflect intuition, the degeneracy between parameters is artificially destroyed.

@@

Once determined, this average covariance matrix could be delivered to the communities that created the synthetic libraries, which would enable them to rapidly pinpoint and correct non-physical lines in the synthetic models. Alternatively, we could correct the models ourselves by using the chain of logic and mathematical post-processing that we used to created the synthetic model spectrum to reverse-engineer what the behavior of the raw synthetic spectrum \emph{should} be, at the raw $R \gtrsim$~100,000 resolution. This fundamental application of machine learning would enable us to create our own library of data driven, semi-empirical stellar models. Rather than simply assembling an empirical spectral library using only real stellar spectra, this combined approach is more powerful because the synthetic stellar atmospheres provide an actual anchor point of fundamental stellar parameters tied down by the laws of stellar physics.

\section{Examples}
\label{sec:examples}

@@

To illustrate the performance of our technique, we use two datasets with varying properties. The first is a high-resolution optical spectrum of an F-star, WASP-14. The second is a moderate resolution, mid-infrared spectrum of Gl 51.

WASP-14 is a metal-poor F star acquired with the \emph{TRES} instrument \citep{mink11} on the 1.5m telescope at FLWO. The data was previously analyzed in the context of transiting exoplanet hosts by \citet{torres12}. Automated reduction pipeline by Buchhave. Has XX S/N per resolution element in the \ion{Mg}{b} order ($\sim 5100$ \AA). 51 echelle orders spanning from XX to XX.

Gl 51 is a M5 star from the IRTF spectral library \citep{cushing05, rayner09} taken with the SPEX instrument \citep{rayner03}. CITE (Rojas-Ayala) use the H20 -K index, defined over what wl range? We do this as well as more general tests.

The SPC analysis within \citet{torres12} focused on a region of spectrum from $5000 5300$ \AA and used many methods. The other methods are not directly comparable because they use different underlying assumptions. Additionally, there are systematics introduced by metallicity assumptions and radiative transfer. Using this same spectral library as SPC, we recover the same parameters.

Describe the instruments. Describe when it was taken. Describe the noise properties, etc. Maybe a table might be worthwhile to describe the spectra. Here also describe the WASP-14 fit using the Kurucz spectra, using a fixed log g, and state that they are the same as Willie's paper. Talk about putting things in a Bayesian framework, using a sensitivity to a prior, to a $\delta$-function prior. Point out that this Bayesian approach allows you to do this.

@@

Talk about giving more reasonable errors using regions, in the biased case.
Show a figure of the regions determined, plotted with the increased variance envelope over the residuals.
